---
title: "week2 project"
author: "hongyuatcufe"
date: "2022/2/2"
output:
  html_document: default
---

# Week2 project by hongyuatcufe

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.Loading and preprocessing the data

Read activity.csv file to variable activity.


```{r echo=TRUE}
activity=read.csv('activity.csv')
activity[[2]]=as.Date(activity[[2]])
head(activity)
```

## 2.Caculate the mean total number of steps taken per day

Calculate the total number of steps taken per day

```{r echo=TRUE,message=FALSE}
library(dplyr)

steps_per_day <- activity %>%
  group_by(date) %>%
  summarise(steps=sum(steps,na.rm=TRUE))

head(steps_per_day)
```

Make a histogram of the total number of steps taken each day

```{r echo=TRUE}
# set breaks to 10 to make the histogram show more details
hist(steps_per_day$steps,breaks=10)
```

Calculate and report the mean and median of the total number of steps taken per day

```{r echo=TRUE}
# Caculate the mean and median of steps
steps_mean=mean(steps_per_day$steps)
steps_median=median(steps_per_day$steps)
# print the results
print(paste('The mean of steps per day is',round(steps_mean,2),'and',
      paste("The median of steps per day is",steps_median)))
```

## 3.Caculate the average daily activity pattern

time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo=TRUE,message=FALSE}
steps_per_interval <- activity %>%
  group_by(interval) %>%
  summarise(steps=mean(steps,na.rm=TRUE))

plot(steps_per_interval,type='l')
max_step_interval=steps_per_interval[steps_per_interval$steps==max(steps_per_interval$steps),'interval']
print(paste('The max of the average number of steps taken occur in interval', max_step_interval))
```

The max of the average number of steps taken occur at 8:35 AM. 

## Imputing missing values



