---
title: "week2 project"
author: "hongyuatcufe"
date: "2022/2/5"
output:
  html_document: default
  pdf_document: default
---

# Week2 project by hongyuatcufe

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_knit$set(base.dir = 'figs')
```

## 1.Loading and preprocessing the data

Read activity.csv file to variable activity.


```{r echo=TRUE}
activity=read.csv('activity.csv')
activity[[2]]=as.Date(activity[[2]])
```

## 2.Caculate the mean total number of steps taken per day

Calculate the total number of steps taken per day

```{r echo=TRUE,message=FALSE}
library(dplyr)

steps_per_day <- activity %>%
  group_by(date) %>%
  summarise(steps=sum(steps,na.rm=TRUE))

head(steps_per_day)
```

Make a histogram of the total number of steps taken each day

```{r echo=TRUE}
# set breaks to 10 to make the histogram show more details
hist(steps_per_day$steps,breaks=10)
```

Calculate and report the mean and median of the total number of steps taken per day

```{r echo=TRUE}
# Caculate the mean and median of steps
steps_mean=mean(steps_per_day$steps)
steps_median=median(steps_per_day$steps)
# print the results
print(paste('The mean of steps per day is',round(steps_mean,2),'and',
      paste("The median of steps per day is",steps_median)))
```

## 3.Caculate the average daily activity pattern

time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo=TRUE,message=FALSE}
steps_per_interval <- activity %>%
  group_by(interval) %>%
  summarise(steps=mean(steps,na.rm=TRUE))

plot(steps_per_interval,type='l')
max_step_interval=steps_per_interval[steps_per_interval$steps==max(steps_per_interval$steps),'interval']

print(paste('The max of the average number of steps taken occur in interval', max_step_interval,'and the max steps is about',round(max(steps_per_interval$steps),0),'steps!'))
```

The max of the average number of steps taken occur at 8:35 AM. 

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r echo=TRUE}

steps_missing=sum(is.na(activity$steps))
print(paste('The total number of missing values in the dataset is',steps_missing))

```

filling in all of the missing values in the dataset with mean of steps of relevant interval and Create a new dataset

```{r echo=TRUE}
activity_changed = activity %>%
  left_join(.,steps_per_interval,by="interval") %>%
  mutate(steps.x=ifelse(is.na(steps.x),steps.y,steps.x)) %>%
  select(steps.x,date,interval) %>%
  rename(steps=steps.x)

# the NA in dataset activity changed to mean for relevant 5-minute interval as blow:
head(activity_changed)
```


Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

```{r echo=TRUE}

steps_per_day_new <- activity_changed %>%
  group_by(date) %>%
  summarise(steps=sum(steps))

hist(steps_per_day_new$steps,breaks=10)

# Caculate the mean and median of steps
steps_mean_new=mean(steps_per_day_new$steps)
steps_median_new=median(steps_per_day_new$steps)
# print the results
print(paste('The mean of steps per day  with missing values filled is',round(steps_mean_new,2),'and',
      paste("The median of steps per day with missing values filled is",round(steps_median_new,2))))

```


After missing values in the dataset is imputed by the mean of steps number of relevent interval, the mean step increased from 9354.23 to 10766.19, and median increased from 10395 to 10766.19. It's worth noting that After imputing, the mean and median of steps become the same.


## Differences in activity patterns between weekdays and weekends

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r echo=TRUE}
activity_changed = activity_changed %>%
  mutate(is_weekday=case_when(weekdays(date) %in% c("星期一",
                                                    "星期二",
                                                    "星期三",
                                                    "星期四",
                                                    "星期五") ~ 'weekday',
                              weekdays(date) %in% c("星期六",
                                                    "星期日") ~ 'weekend'))

table(activity_changed$is_weekday)

steps_per_interval_changed <- activity_changed %>%
  group_by(interval,is_weekday) %>%
  summarise(steps=mean(steps,na.rm=TRUE))

head(steps_per_interval_changed)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r echo=TRUE}
library(ggplot2)

steps_per_interval_changed %>%
  ggplot( aes(x=interval, y=steps)) +
  geom_line() +
  facet_wrap(~is_weekday,nrow=2)

```


A better way to compare the step of weekday and weekend is to put two chart together and show lines in different color as blow:

```{r echo=TRUE}
steps_per_interval_changed %>%
  ggplot( aes(x=interval, y=steps, group=is_weekday,color=is_weekday)) +
  geom_line()

```


It seems that steps number in 8:00 AM to 9:00 AM of weekend is less than weekday, and steps number of weekend is much more than weekday in 11:00 AM to 12:00 AM and 4:00 PM to 6:00 PM. 
